### 2018-02-11 dataset.tex
final round of edits--hopefully--before submission

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-02}
    \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between simulations for three different velocity models and data recorded at ground motion monitoring stations. The color version of this figure is available only in the electronic edition.}
    \label{fig:ref-gof-maps}
\end{figure*}
% 
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\columnwidth]{figures/pdf/figure-03}
    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component of motion (NS, EW, UD). In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots. The color version of this figure is available only in the electronic edition.}
    \label{fig:data-box-plot}
\end{figure}

>>>>

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-02}
    \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between simulations for three different velocity models and data recorded at ground motion monitoring stations. The color version of this figure is available only in the electronic edition.}
    \label{fig:ref-gof-maps}
\end{figure*}
% 
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\columnwidth]{figures/pdf/figure-03}
    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component of motion (NS, EW, UD). In each case, the median is indicated by a notch in the box, the boxes represents the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), and the line shows the range of the data, with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots. The color version of this figure is available only in the electronic edition.}
    \label{fig:data-box-plot}
\end{figure}

### 2018-02-11 results.tex
finished first complete version
som edits before submission
final round of edits--hopefully--before submission

Some aspects of Fig.~\ref{fig:clusters} are worth discussing. Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case. In the C5-C9 and C7-C9 combinations in the 3-feature analysis, for instance, the excellent cluster has a limited presence. This is due to the influence of the lower C9 values (see Fig.~\ref{fig:data-box-plot}) on the correlations with C5 and C7. This is not necessarily a bad thing, as it helps to understand the different relationships between the various GOF metrics. In the best of cases, the combinations C1-C2 and C5-C8 in the 2- and 4-feature analysis, for instance, exhibit an almost perfect proportionality between the metrics. That means either one of the metrics in those combinations is redundant. Relationships with redundant features are those where knowledge about one of the features provides a direct view into the other. On the other hand, the combination C1-C7 in the 2-feature analysis presents an example of an irrelevant feature. We say C1 is irrelevant because it provides no insight about the outcome of C7. Identifying subspaces with redundant and irrelevant features is important because, on the one hand, the former help reduce the number of necessary features, while on the other hand, the latter can essentially be discarded because of their weak contribution to the decision making process \citep{Dy_2004_MLR}.

@@

@@

\begin{figure*}[t]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
	\caption{Accuracy of tree predictions in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed with respect to the number of attributes (GOF metrics) as a function of: \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees. The rings around the left-most bottom dots indicate selected trees shown in Fig.~\ref{fig:trees}. The color version of this figure is available only in the electronic edition.}
	\label{fig:nodes-depth}
\end{figure*}

@@

\begin{table}%[t]
	\centering
	\caption{Confusion matrix results for decision tree T1}
	\label{tab:t1:confusion}
	\small
	\begin{tabular}{cccccc}
		& 		& \multicolumn{4}{c}{Prediction}\\ 
		\cline{3-6}
		&   	& P  	& F     & G  	& E 	\\
		\cline{2-6}
		\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{Actual}}}  
		& P 	& 242 	& 11 	& 0     & 0    	\\
		& F 	& 13 	& 313   & 16    & 0    	\\
		& G 	& 0 	& 42    & 207   & 15   	\\
		& E 	& 0 	& 0     & 23    & 231	\\ 
		\cline{2-6}
	\end{tabular}
\end{table}

\begin{table}%[t]
	\centering
	\caption{Confusion matrix results for decision tree T3}
	\label{tab:t3:confusion}
	\small
	\begin{tabular}{cccccc}
		& 		& \multicolumn{4}{c}{Prediction}\\ 
		\cline{3-6}
		&   	& P  	& F     & G  	& E 	\\
		\cline{2-6}
		\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{Actual}}}  
		& P 	& 248 	& 14 	& 0     & 0    	\\
		& F 	& 7 	& 343   & 18    & 0    	\\
		& G 	& 0 	& 9 	& 221   & 33   	\\
		& E 	& 0 	& 0     & 7    	& 213	\\ 
		\cline{2-6}
	\end{tabular}
\end{table}

\begin{table}
	\centering
	\caption{T1 precision, recall, and $F_1$ values per class}
	\label{tab:t1:f1}
	\small
	\begin{tabular}{lccc}
	Class   & $P$	& $R$  	& $F_1$ \\ 
	\hline
	P 		& 0.96 	& 0.95 	& 0.95 	\\
	F 		& 0.92 	& 0.86 	& 0.88 	\\
	G 		& 0.78 	& 0.84 	& 0.81 	\\
	E 		& 0.91 	& 0.94 	& 0.92 	\\
	\hline
	Mean 	&		&		& 0.90	\\
	\end{tabular}
\end{table}

\begin{table}
	\centering
	\caption{T3 precision, recall, and $F_1$ values per class}
	\label{tab:t3:f1}
	\small
	\begin{tabular}{lccc}
	Class   & $P$	& $R$  	& $F_1$ \\ 
	\hline
	P 		& 0.95 	& 0.97 	& 0.95 	\\
	F 		& 0.93 	& 0.93 	& 0.93 	\\
	G 		& 0.84 	& 0.89 	& 0.86 	\\
	E 		& 0.96 	& 0.86 	& 0.91 	\\
	\hline
	Mean 	&		&		& 0.91	\\
	\end{tabular}
\end{table}

@@

The final selection of a preferred tree, then, comes down to reducing the complexity. We favor tree T1 because it is based only in three decision steps, and two GOF metrics, C4: energy, and C8: response spectra.

>>>>

Some aspects of Fig.~\ref{fig:clusters} are worth discussing. Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case. In the C5-C9 and C7-C9 combinations in the 3-feature analysis, for instance, the excellent cluster has a limited presence. This is due to the influence of the lower C9 values (see Fig.~\ref{fig:data-box-plot}) on the correlations with C5 and C7. This is not necessarily a bad thing, as it helps to understand the different relationships between the various GOF metrics. In the best of cases, the combinations C1-C2 and C5-C8 in the 2- and 4-feature analysis, for instance, exhibit an almost perfect proportionality between the metrics. That means either one of the metrics in those combinations is redundant. Relationships with redundant features are those where knowledge about one of the features provides a direct view into the other. On the other hand, the combination C1-C7 in the 2-feature analysis presents an example of an irrelevant feature. We say C1 is irrelevant because it provides no insight about the outcome of C7 or that of the clusters. Identifying subspaces with redundant and irrelevant features is important because, on the one hand, the former help reduce the number of necessary features, while on the other hand, the latter can essentially be discarded because of their weak contribution to the decision making process \citep{Dy_2004_MLR}.

@@

\begin{figure*}[t]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
	\caption{Accuracy of tree predictions in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed with respect to the number of attributes (GOF metrics) as a function of: \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees. The rings around some of the dots indicate select trees used as reference in the text, other figures and tables. The color version of this figure is available only in the electronic edition.}
	\label{fig:nodes-depth}
\end{figure*}

\begin{table}[t]
	\centering
	\caption{Confusion matrix results for decision tree T1}
	\label{tab:t1:confusion}
	\small
	\begin{tabular}{cccccc}
		& 		& \multicolumn{4}{c}{Prediction}\\ 
		\cline{3-6}
		&   	& P  	& F     & G  	& E 	\\
		\cline{2-6}
		\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{Actual}}}  
		& P 	& 242 	& 11 	& 0     & 0    	\\
		& F 	& 13 	& 313   & 16    & 0    	\\
		& G 	& 0 	& 42    & 207   & 15   	\\
		& E 	& 0 	& 0     & 23    & 231	\\ 
		\cline{2-6}
	\end{tabular}
\end{table}

\begin{table}[t]
	\centering
	\caption{Confusion matrix results for decision tree T3}
	\label{tab:t3:confusion}
	\small
	\begin{tabular}{cccccc}
		& 		& \multicolumn{4}{c}{Prediction}\\ 
		\cline{3-6}
		&   	& P  	& F     & G  	& E 	\\
		\cline{2-6}
		\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{Actual}}}  
		& P 	& 248 	& 14 	& 0     & 0    	\\
		& F 	& 7 	& 343   & 18    & 0    	\\
		& G 	& 0 	& 9 	& 221   & 33   	\\
		& E 	& 0 	& 0     & 7    	& 213	\\ 
		\cline{2-6}
	\end{tabular}
\end{table}

@@

@@

\begin{table}[t]
	\centering
	\caption{T1 precision, recall, and $F_1$ values per class}
	\label{tab:t1:f1}
	\small
	\begin{tabular}{lccc}
	Class   & $P$	& $R$  	& $F_1$ \\ 
	\hline
	P 		& 0.96 	& 0.95 	& 0.95 	\\
	F 		& 0.92 	& 0.86 	& 0.88 	\\
	G 		& 0.78 	& 0.84 	& 0.81 	\\
	E 		& 0.91 	& 0.94 	& 0.92 	\\
	\hline
	Mean 	&		&		& 0.90	\\
	\end{tabular}
\end{table}

\begin{table}[t]
	\centering
	\caption{T3 precision, recall, and $F_1$ values per class}
	\label{tab:t3:f1}
	\small
	\begin{tabular}{lccc}
	Class   & $P$	& $R$  	& $F_1$ \\ 
	\hline
	P 		& 0.95 	& 0.97 	& 0.95 	\\
	F 		& 0.93 	& 0.93 	& 0.93 	\\
	G 		& 0.84 	& 0.89 	& 0.86 	\\
	E 		& 0.96 	& 0.86 	& 0.91 	\\
	\hline
	Mean 	&		&		& 0.91	\\
	\end{tabular}
\end{table}

@@

Another aspect of interest is the level of participation of each metric in the analysis of data as GOF values are run through the decision nodes in the trees. The results of such level of participation are listed in Table \ref{tab:participation} for trees T1, T2, T3, and T66. (The tree T66 is the most complex of them all as inferred from the number of metrics involved and the number of nodes and levels as seen in Fig.~\ref{fig:nodes-depth}.) The percentages in this table represent the amount of data that is seen by a decision node associated by any particular metric. The nodes are not unique and metrics are often used in different nodes in a given tree, thus these percentage are cumulative for each individual metric. A percentage of 100 means a given metric has the opportunity to see all data samples through a single nodes or multiple ones. A low percentage means only a few data samples are evaluated by that metric, and a null percentage means the metric plays no role in the decision tree (it is not present in any node).

Notice that almost independently of the tree (T3 being the exception), the metrics C4 (total energy) and C8 (response spectra) are consistently the most relevant ones in the decision algorithms. They are followed by C5 (peak acceleration) and C6 (peak velocity). On the other end, notice that C11 (strong phase duration) plays no role whatsoever in any of the trees, and that C1 (Arias integral), C2 (energy integral), and C3 (Arias intensity) have only small to marginal participations. The remaining metrics---C7 (peak displacement), C9 (Fourier spectrum), and C10 (Cross correlation)---have a somewhat significant participation, but only in the more complex trees.

A low participation does not necessarily mean that a given metric is not important. Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering. What the results we present here mean, in the context of GOF validation, is that in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more decisive in reaching an accurate conclusion about validation quality than the duration (C11).

In the end, the final selection of a preferred tree comes down to reducing complexity in the analysis without compromising the interpretation of results. We favor tree T1 because it is based only in three decision steps, and two GOF metrics: C4 (total energy), and C8 (response spectra).

\begin{table}[t]
	\centering
	\caption{Metrics data analysis participation in select trees (in percent)}
	\label{tab:participation}
	\small
	\begin{tabular}{lrrrr}
	Metric 	& T1 	& T2	& T3 	& T66 	\\ 
	\hline
	C1 		& --	& --	& -- 	& 1.5	\\
	C2 		& -- 	& --	& -- 	& 0.5	\\
	C3 		& -- 	& -- 	& -- 	& 9.6	\\
	\rowcolor{light-light-gray}
	C4 		& 46.8	& 52.4	& 60.2	& 100.0	\\
	\rowcolor{light-light-gray}
	C5 		& -- 	& 21.5	& 21.5	& 72.2	\\
	\rowcolor{light-light-gray}
	C6 		& -- 	& -- 	& 37.0	& 56.3	\\
	C7 		& -- 	& -- 	& -- 	& 19.4	\\
	\rowcolor{light-light-gray}
	C8 		& 100.0	& 100.0	& 100.0	& 100.0	\\
	C9 		& -- 	& -- 	& --	& 23.8	\\
	C10		& -- 	& -- 	& -- 	& 22.2	\\
	C11		& -- 	& -- 	& -- 	& -- 	\\
	\hline
	\end{tabular}
\end{table}
